---
title: "MKT 552 Market Basket Analysis Assignment"
author: "Aridania Gerardo"
date: "2022-07-30"
output: html_document
---

## Installing Packages

```{r}
#packages for analysis
library(dplyr)
library(readr)
library(arules)
library(arulesViz)
library(scales)
```

## Uploading data

```{r}
Order<-read.csv("https://raw.githubusercontent.com/aridaniaG/InstacartMKTBasket/main/Basket_Raw.csv", nrows=100000)
Products<-read.csv("https://raw.githubusercontent.com/aridaniaG/InstacartMKTBasket/main/products.csv")
DepartmentName<-read.csv("https://raw.githubusercontent.com/aridaniaG/InstacartMKTBasket/main/departments.csv")
```

## Preprocessing data

```{r}
#extracting columns for df
OrderId<-Order[,1:2]
ProductDepartmentName<-Products[,c(1:2,4)]
```

```{r}
#transforming df to transactions for analysis
ProductDepartmentName<-merge(ProductDepartmentName, DepartmentName, by="department_id")
TransData<-merge(OrderId, ProductDepartmentName, by="product_id")
TransData<-arrange(TransData, OrderId)
```

```{r}
#data for analysis
ProductAnaly<-TransData[,c(2,4)]
DepartmentAnaly<-TransData[,c(2,5)]
```

```{r}
#data for both department and specific products
TransPData <- split(ProductAnaly$product_name, ProductAnaly$order_id)
TransDData <- split(DepartmentAnaly$department, DepartmentAnaly$order_id)
```

```{r}
ProTran = as(TransPData,"transactions")
DepTran = as(TransDData,"transactions")
```

## EDA

### Specific Products

The five most frequent products bought on instacart are Bananas, Bag of Organic Bananas, Organic Strawberries, Organic Baby Spinach, and Large Lemon's.

```{r}
#frequent items for specific prodcuts
frequentItems <- eclat(ProTran, parameter = list(supp = 0.07, maxlen = 15))
inspect(frequentItems)
#plot frequency
itemFrequencyPlot(ProTran, topN=5, type="absolute", main="Item Frequency per Product")
```

Organic Hass Avocado, Organic Raspberries, and Organic Unsweetened Almond Milk are most likely to be purchased with a Bag of Organic Bananas with a confidence level of 0.83.

```{r}
#rules for Products
rules <- apriori(ProTran, parameter = list(supp = 0.001, conf = 0.5, maxlen=15))
#confidence
rules_conf <- sort(rules, by="confidence", decreasing=TRUE)
inspect(head(rules_conf))
```

### Departments

The top five most popular categories that are bought from are produce, dairy eggs, beverages, snacks, and frozen items.

```{r}
#frequent items in departments
frequentItems <- eclat(DepTran, parameter = list(supp = 0.07, maxlen = 15))
inspect(frequentItems)
#plot frequency
itemFrequencyPlot(DepTran, topN=5, type="absolute", main="Item Frequency per Categories")
```

Purchase of dairy eggs is associated with bulk, personal care, and snacks. Purchase of produce is associated with bulk, dairy eggs, personal care.

```{r}
#rules for departments
rules <- apriori(DepTran, parameter = list(supp = 0.001, conf = 0.5, maxlen=15))
#confidence
rules_conf <- sort(rules, by="confidence", decreasing=TRUE)
inspect(head(rules_conf))
```

## Analysis Summary

Items that are likely to be bought together are from the produce department such as dairy eggs and produce with snacks and personal care.

If consumers buy two products together, the leader of the basket is most likely to be an item from the produce department sand the add-on is most likely to be itmes that would accompany such as beverages and snacks.
